---
title: "Untitled"
author: "Group02"
number-sections: false
format: 
  html:
    embed-resources: true
    code-tools: true
  pdf: default
editor_options: 
  chunk_output_type: console
editor: visual
execute:
  warning: false
  message: false
output:
  pdf_document:
---

## exploratory data analysis
#hello world
We are using the given data set "dataset02", which provides data on family income and expenditure. Our aim is to analyse which household related variables influence the number of people living in a household?

```{r}
# Load package
library(tidyverse)
library(moderndive)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)
```

```{r}
# Upload the dataset02 and rename it as "data"
data <- read.csv("dataset02.csv")
glimpse(data)
```

```{r}
sapply(data, function(x) if(is.factor(x)) levels(x))
ggplot(data, aes(Total.Number.of.Family.members)) +
  geom_histogram()
```

Converting character variables into factor variables is convenient for analysis. We find that "Region" only have one level, so we won't use this variables to fit model.

The level of "Household.Head.Sex" is female and male. "Type.of.Household" has three different level, which are extended Family, single family and two or more nonrelated persons/members. Because of "1" of variable "Electricity" means the house have electricity and "0" means the house doesn't have electricity, we convert the count variable "Electricity" into a factor variable.

```{r}
# Convert some variables into factor variables
data$Region <- as.factor(data$Region)
data$Household.Head.Sex <- as.factor(data$Household.Head.Sex)
data$Type.of.Household <- as.factor(data$Type.of.Household)
data$Electricity <- as.factor(data$Electricity)
```

```{r}

# View variable levels
str(data)
levels(data$Region)
levels(data$Household.Head.Sex)
levels(data$Type.of.Household)
levels(data$Electricity)
```

## Methodology

The response variable here is the number of people living in the household (Total.Number.of.Family.members), which is count data. So we consider using Poisson regression，which is suitable for count data.

### Poisson model

We plan to select model by looking at the p-value of variable and using the stepwise model selection based on AIC. Model 1 covers all explanatory variables.

```{r}
poisson_model_1 <- glm(Total.Number.of.Family.members ~ Total.Household.Income +
                         Total.Food.Expenditure + 
                         Household.Head.Sex + 
                         Household.Head.Age + 
                         Type.of.Household + 
                         House.Floor.Area + 
                         House.Age + 
                         Number.of.bedrooms + 
                         Electricity,
                       family = "poisson", data = data)

summ(poisson_model_1)
plot(poisson_model_1)
```

From the result, we notice that P-value of house floor area is 0.77, indicating that it is not statistically significant. So we decide to eliminate this variable.

```{r}
poisson_model_2 <- glm(Total.Number.of.Family.members ~ Total.Household.Income +
                         Total.Food.Expenditure + 
                         Household.Head.Sex +
                         Household.Head.Age + 
                         Type.of.Household + 
                         House.Age + 
                         Number.of.bedrooms + 
                         Electricity,
                       family = "poisson", data = data)
summ(poisson_model_2)
```

The next step is remove the insignificant electricity variables.

```{r}
poisson_model_3 <- glm(Total.Number.of.Family.members ~ Total.Household.Income +
                         Total.Food.Expenditure + 
                         Household.Head.Sex + 
                         Household.Head.Age + 
                         Type.of.Household + 
                         House.Age + 
                         Number.of.bedrooms,
                       family = "poisson", data = data)
summ(poisson_model_3)
```

```{r}
poisson_model_4 <- glm(Total.Number.of.Family.members ~ Total.Household.Income +
                         Total.Food.Expenditure + 
                         Household.Head.Sex + 
                         Household.Head.Age + 
                         Type.of.Household + 
                         House.Age,
                       family = "poisson", data = data)
summ(poisson_model_4)
#plot(poisson_model_4)
```
